\section{Introduction} \label{sec:introduction}
Topological analysis methods can be applied on scientific data to generate
multi-resolution hierarchies that encode features of interest in the data.
The analysis methods discretize the domain into sets of features 
which we denote feature families, that can be split and merged according to
an importance measure.  In general, analysis methods based on Reeb graphs 
and their variants, or the Morse Smale (MS) complex generate similar outputs: 
a segmentation of the input domain and a one-parameter simplification sequence 
for coarsening or refining the segmentation.  The Reeb graph segments the domain 
based on the evolution of the level sets of a function $f$, while the Morse Smale 
complex segments the domain based on the gradient flow of $f$.  The simplification 
sequence can be driven by (1) traditional persistence, where the importance of a 
feature is defined as the difference in function value between its birth and 
death or (2) threshold based simplification, where all features above or below 
a particular threshold are disregarded.

Scientific data is increasingly high-dimensional with scientists tracking large
numbers of field attributes within their simulations and experiments.  Most
topological analysis techniques segment the data according to one of these
field attributes.  In an effort to expand the general applicability of 
toplogical analysis tools, we are creating a unified semantic framework based on
the notion of feature families that (1) allows for in depth analysis of the 
remaining field parameters within and across segmented features, (2) facilitates 
the development of downstream analysis tools, and (3) compresses
the data to be stored, while maintaining the information that is necessary for 
additional analyses downstream.  We are developig a topological file format that 
provides a general interface to topological codes by storing information on each 
feature, namely a unique id, a list of domain elements that comprise the feature, 
a list of feature properties (size, mean function values, etc.), and persistence
information to encode birth, death, and merge events.

This document describes our efforts to develop this framework and is outlined as 
follows:  the notion of feature family processing is discussed in section 
\ref{sec:featureFamilyProcessing} and the  topology file format is presented in 
section \ref{sec:topologyFileFormat}.  Details for installing and using existing 
downstream tools are included in sections \ref{sec:installation} and \ref{sec:usage} 
respectively.

